\documentclass[AMA,STIX1COL,APA,STIX2COL]{WileyNJD-v2}


% Pandoc citation processing
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
% for Pandoc 2.8 to 2.10.1
\newenvironment{cslreferences}%
  {}%
  {\par}
% For Pandoc 2.11+
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc} % for calculating minipage widths
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{xcolor}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{colortbl}
\usepackage{hhline}
\usepackage{longtable}
\usepackage{array}
\usepackage{hyperref}

\articletype{Research article}

\received{2017-01-01}

\revised{2017-02-01}

\accepted{2017-03-01}

\raggedbottom

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\hypersetup{
    colorlinks = true,
    linkbordercolor = {white},
    linkcolor=blue
}

\usepackage{lineno} % add
  \linenumbers % turns line numbering on

\begin{document}

\title{Exploring team roles for social innovation labs: toward a
competence-based self-assessment approach}

\author[ERPI,UN]{Ferney Osorio*}
\author[ERPI]{Fabio A. Cruz Sanchez}
\author[ERPI]{Laurent Dupont}
\author[ERPI]{Mauricio Camargo}

\authormark{Uthor \emph{et al}.}

\address[ERPI]{Université de Lorraine - ERPI -F-54000, Nancy, France,}
\address[UN]{Departamento de Ingeniería de Sistemas e Industrial,
Universidad Nacional de Colombia, Bogotá, Colombia}

\corres{*Corresponding author name, This is sample corresponding
address. \email{authorone@gmail.com}}

\presentaddress{This is sample for present address text this is sample
for present address text}

\abstract{Recently, there has been a great interest in the development
of innovations labs as intermediate means to fostering social innovative
solutions to wicked problems. However, understanding how lab teams are
assembled including the underlying competences and main roles inside of
these organizational structures is still yet to be addressed. This
aspect is of paramount importance at the early-design phase to foster
the future development and consolidation of such initiatives. A
competence-based role model is proposed as a basis for guiding the
conformation of social innovation lab teams. The model has been
structured from (1) a set of 14 competences for social innovation labs
retrieved from the literature, (2) a comparison of 7 frameworks of
innovation team roles and (3) authors experience. The proposed model is
then operationalized through a self-assessment approach composed of an
online questionnaire and a retrospective workshop aiming to allow team
members to position themselves in terms of the potential role that they
could perform for their team but also to elicit improvement strategies.
The self-assessment methodology is then applied among 10 Latin American
nascent social innovation lab teams with focus on climate change
challenges. Insights and implications of this exploratory study for both
researchers and practitioners are then discussed.}

\keywords{innovation lab, social innovation, innovation team roles,
innovation competences, self-assessment}

\maketitle

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Today's most critical challenges demand systemic ways to tackle them.
Climate change, environmental degradation, health crisis, education
inequalities, and employment and poverty reduction are some examples of
those wicked problems characterized by their complexity, their
interdependencies and their context specificity
(\protect\hyperlink{ref-Zivkovic2018}{Zivkovic 2018}). Social Innovation
(SI) then emerges as a research strand for not only helping to
understand these societal issues but to facilitate the development of
systemic strategies toward a transformative change of social practices
in order to solve social problems and meeting local demands
(\protect\hyperlink{ref-Strasser2019}{Strasser, Kraker, and Kemp 2019}).

In particular, the notion of SI Labs has recently become a subject of
interest in the literature. SI labs emerge as an approach to keep up
with increasing changes and accumulating challenges that society deals
with and where more conventional approaches relying solely on
techno-centric approaches fall short
(\protect\hyperlink{ref-Westley2015}{Westley et al. 2015};
\protect\hyperlink{ref-Jezierski2014}{Jezierski et al. 2014}).
Innovation labs are defined as semi-autonomous organizations dedicated
to facilitate innovation processes by allowing multi-stakeholders groups
to interact in open collaboration with the purpose of creating and
prototyping solutions to systemic challenges while strengthening
people's innovative and technological competences
(\protect\hyperlink{ref-Lewis2005}{M. Lewis and Moultrie 2005};
\protect\hyperlink{ref-Gryszkiewicz2016}{Gryszkiewicz, Lykourentzou, and
Toivonen 2016}; \protect\hyperlink{ref-Zivkovic2018}{Zivkovic 2018}). In
this sense, SI labs act as cross-pollinators of knowledge, creating
dialogue, mixing voices, allowing for new ideas to appear and to be
translated into alternative solutions
(\protect\hyperlink{ref-Wascher2019}{Wascher, Kaletka, and Schultze
2019}). The way these organizational forms perform often depends on the
problem that is being addressed and the context they belong to. This
means that the people, the organizations involved and even the methods
applied within a SI lab are constantly changing
(\protect\hyperlink{ref-Wascher2018}{Wascher et al. 2018}).

The changing and permeable nature of the ``lab'' phenomenon represents a
complex working environment which often leads to conditions of
uncertainty. This is something on which authors have raised concerns,
suggesting that those teams in charge of leading an innovation lab
should be able to deal with ambiguities, integrate multiple perspectives
and facilitate the work across-disciplines
(\protect\hyperlink{ref-Osorio2020}{Osorio et al. 2020};
\protect\hyperlink{ref-McGann2019}{McGann, Wells, and Blomkamp 2019}).
There is an increasing interest in how innovation labs can be used to
address societal problems (\protect\hyperlink{ref-McGann2019}{McGann,
Wells, and Blomkamp 2019}); governments, companies, universities, and
even communities are continuously turning to the implementation of their
own ``lab.'' They are becoming vectors for fostering collaborative
learning, inclusive entrepreneurial thinking, systemic change and the
transfer of innovation capabilities
(\protect\hyperlink{ref-Delgado2020}{Delgado et al. 2020};
\protect\hyperlink{ref-RezaeeVessal2021}{Vessal et al. 2021};
\protect\hyperlink{ref-Rayna2019}{Rayna and Striukova 2019};
\protect\hyperlink{ref-Camargo2021}{Camargo, Morel, and Lhoste 2021}).
However, several questions arise when considering how the teams managing
these initiatives should be composed and organized
(\protect\hyperlink{ref-Lewis2020}{J. M. Lewis 2020};
\protect\hyperlink{ref-Zivkovic2018}{Zivkovic 2018}). This aspect is of
paramount importance, especially at the early-design phase to foster the
future development and consolidation of such initiatives.

In fact, the assembling of innovation teams has been a matter of
interest for a long time. Both practitioners and scholars from public
and private sectors have addressed and shared their experiences in the
nature and characteristics of innovations teams across time. This is a
vision that has been in constant evolution, referring for instance to
the 80's where corporate-type innovation teams, whose nature depended
mainly on the emergence of those `champions' capable of overcoming any
obstacle, while additional roles were organized towards supporting them
(\protect\hyperlink{ref-Roberts1982}{Roberts and Fusfeld 1982};
\protect\hyperlink{ref-Jenssen2004}{Jenssen and Jørgensen 2004}). Then,
as the adoption of open innovation practices became more widespread, the
idea of innovation teams has progressively become more agile and
adaptive, opening the door to the integration of multiple disciplines,
and being the inspiration for new ways of work and collaboration
(\protect\hyperlink{ref-Hellstrom2002}{Hellström, Jacob, and Malmquist
2002}; \protect\hyperlink{ref-Hering2005}{Hering and Phillips 2005};
\protect\hyperlink{ref-Gemunden2007}{Gemünden, Salomo, and Hölzle
2007}). While the perception of lonely innovators and isolated teams
seems to persist today, the increasing interconnection and complexity of
the problems that we face as society, the amount of information and
knowledge that is continuously created, and the challenging task of
making critical decisions with unforeseeable repercussions, are evidence
on why today's innovation teams are called to be able to efficiently
collaborate under a multitude of perspectives, disciplines and cultures
(\protect\hyperlink{ref-Bjorklund2017}{Björklund et al. 2017};
\protect\hyperlink{ref-Puttick2014-Teams}{Puttick, Baeck, and Colligan
2014}).

This is not a minor issue since it is in people where the success of
every innovation process of an organization lies
(\protect\hyperlink{ref-Leonard1995}{Leonard-Barton 1995}). Thus,
understanding the dynamics of group work and team performance has been a
topic of interest for the scientific community. In this regard, previous
studies have tackled this issue from several perspectives. In terms of
team theory for instance, \protect\hyperlink{ref-Belbin2010}{Belbin}
(\protect\hyperlink{ref-Belbin2010}{2010}) gathers in her book an
extensive research that resumes her proposition of the nine key team
roles at work. Originally published in 1993, Belbin explains in her work
why roles in a team are in fact the sum up of multiple factors such as
personal traits, knowledge, skills, experience and even situations that
will determine a person's behaviour in group work or in a specific job.
More specifically studies on innovation teams as the ones by
(\protect\hyperlink{ref-Kratzer2004}{Kratzer, Leenders, and Engelen
2004}; \protect\hyperlink{ref-Kratzer2006-A}{Kratzer, Leenders, and
Engelen 2006}; \protect\hyperlink{ref-Kratzer2006-B}{Kratzer, Leenders,
and Engelen 2006}) have focused on examining how factors such as team
communication, conflicts or virtuality influence creativity performance.
Likewise, \protect\hyperlink{ref-DeCusatis2008}{DeCusatis}
(\protect\hyperlink{ref-DeCusatis2008}{2008}) pointed out how team
performance varies based on generational preferences, habits and the
nature of the intended innovation. Precisely, this changing nature of
the innovation process across time has opened the door for not only
asking which roles are required but also what are the competences needed
for successful innovation teams.

That is why researches like these of
(\protect\hyperlink{ref-Chatenier2010}{Chatenier et al. 2010}) and
(\protect\hyperlink{ref-Podmetina2018}{Podmetina et al. 2018}) have
proposed specific competence profiles, for open innovation teams
revealing what are the main tasks they perform, the main challenges they
face and the underlying competencies behind them.
\protect\hyperlink{ref-Chatenier2010}{Chatenier et al.}
(\protect\hyperlink{ref-Chatenier2010}{2010}) pointed out that
competence profiles are instrumental for the creation and development of
innovation teams. However, besides their comprehensive and detailed
model they also suggest that a single competence profile falls short
when it comes to assembling effective innovation teams, specifically at
the moment of determining which competences need to be held by each team
member and for which role.

Despite the existing research, the literature remains scarce when
referring to what competences are key for guiding the conformation of SI
lab teams and under which roles can they be organized. This is a major
issue for the successful implementation of an innovation lab initiative,
since beyond physical and technological resources, human facilitation is
one of its fundamental pillars
(\protect\hyperlink{ref-Magadley2009}{Magadley and Birdi 2009}).
Furthermore, how these aspects are early weighed in terms of the lab
setup and its context (i.e.~private, community or university) determines
the type of challenges a lab team will have to face
(\protect\hyperlink{ref-Rayna2019}{Rayna and Striukova 2019}). But more
importantly, the strategies to overcome these challenges could be driven
or undermined according to the competences of the lab team, reflecting
also on how effectively they would be able to achieve the intended
social impact (\protect\hyperlink{ref-Rayna2021}{Rayna and Striukova
2021}). Therefore, the main focus of this article lies on the
identification of the key competences and roles that could help the
conformation of teams meant to be the bearers of SI processes. Our goal
is to propose a methodological approach for the early design of SI lab
teams. By means of a self-assessment tool, we intend to provide
practical guidance for the creation of more enduring lab teams while at
the same time we continue to create awareness on the management of these
organizational structures.

To this end, the article first elaborates on the concepts of SI lab,
competence and innovation roles. Next, a role-based framework is
developed by comparing seven existing conceptual frameworks drawn from
the literature on innovation teams and SI. Then, the proposed framework
is operationalized through a competence-based assessment tool (online
questionnaire) from which a self-assessment methodology is designed.
This approach is subsequently tested within the context of the Climate
Labs project, an Erasmus+ initiative whose aim is to strengthen the
applied research and innovation capacities of 10 Latin American Higher
Education Institutions in Mexico, Brazil and Colombia via the design and
implementation of Social Innovation Labs for mitigation and adaptation
to Climate Change. Results from this exploratory study evidence that the
chosen approach is instrumental in the characterization of teams at the
early stages of the implementation of a lab project inside Higher
Education Institutions, enabling them to elicit improvement strategies.
Lastly, discussion and conclusions are built around the main
implications of this work and suggested paths for future research.

\hypertarget{theoretical-background}{%
\section{Theoretical Background}\label{theoretical-background}}

\hypertarget{social-innovation-labs}{%
\subsection{Social Innovation Labs}\label{social-innovation-labs}}

SI refers to the new answers provided to the increasing unsatisfied or
badly-satisfied societal issues
(\protect\hyperlink{ref-Gregoire2016}{Gregoire 2016}). It is understood
as the new social relations (doing, organizing, framing and knowing)
between people (e.g.~producers and consumers, citizens and government,
refugees and native inhabitants, etc.) as well as between people and any
other aspect in society (e.g.~people and nature, producers and their
products, etc.) (\protect\hyperlink{ref-Strasser2019}{Strasser, Kraker,
and Kemp 2019}). SI has been described as being context specific, these
new social relations often lead to novel practices that are meant to
address social issues such as childcare, education, unemployment, crime
prevention, ageing population or climate change
(\protect\hyperlink{ref-Rayna2019}{Rayna and Striukova 2019};
\protect\hyperlink{ref-Dias2019}{Dias and Partidário 2019} ). This means
that the value sought through SI is primarily intended to benefit
society rather than individuals
(\protect\hyperlink{ref-Moulaert2014}{Moulaert et al. 2014}). A key
difference from other innovation approaches, such as technological
innovation, is that the focus is not necessarily on new technologies or
material infrastructure but to contribute to solving social problems
where technology is seen as a means for that purpose
(\protect\hyperlink{ref-Mulgan2006}{Mulgan 2006};
\protect\hyperlink{ref-Murray2010}{Murray, Caulier-Grice, and Mulgan
2010}). In that sense, SI mainly consists of taking advantage of
existing competences and expertise within the population to find more
effective, efficient or sustainable ways to tackle current demanding
issues (\protect\hyperlink{ref-Strasser2019}{Strasser, Kraker, and Kemp
2019}). This also implies that SI solutions are a product of relational
changes that prioritize values rather than status, purpose rather than
profit, co-ownership rather than hierarchy, and collaboration rather
than competition (\protect\hyperlink{ref-Strasser2019}{Strasser, Kraker,
and Kemp 2019}; \protect\hyperlink{ref-Gregoire2016}{Gregoire 2016}).

Recently, the term SI lab has been used for framing the different
organizational forms that agglomerate teams and methods with the
intention of creating socially innovative initiatives
(\protect\hyperlink{ref-Jezierski2014}{Jezierski et al. 2014};
\protect\hyperlink{ref-Westley2015}{Westley et al. 2015};
\protect\hyperlink{ref-Wascher2019}{Wascher, Kaletka, and Schultze
2019}). The notion of innovation labs has been present in the literature
for several years now (\protect\hyperlink{ref-Osorio2019-A}{Osorio,
Dupont, Camargo, and Pena 2019}), building on the more classic ``lab''
idea usually associated to the physical or natural sciences, to
establishing itself as a ``safe haven for experimentation, focused on
problem solving and solution creation''
(\protect\hyperlink{ref-Bloom2016}{Bloom and Faulkner 2016}). Among the
constellation of labs, SI labs raise with the particular focus on
addressing complex social problems and enabling coherent action by
multiple stakeholders (\protect\hyperlink{ref-Zivkovic2018}{Zivkovic
2018}). They do so by providing the space and processes for facilitating
collaboration among cross-sector stakeholders in order to develop new
projects, products, tools, regulations, policies, etc.
(\protect\hyperlink{ref-Wascher2018}{Wascher et al. 2018}).

SI labs are characterized because they foster the creation of dialogue,
listening and mixing the different voices of the actors involved, and
creating boundary objects (e.g.~prototypes, illustrations, concepts,
scenarios, and maps) for knowledge co-production processes to allow
diverse actors to work together (See
\protect\hyperlink{ref-Nilsson2015}{Nilsson, Bonnici, and EL 2015};
\protect\hyperlink{ref-Lake2016}{Lake, Fernando, and Eardley 2016};
\protect\hyperlink{ref-Timmermans2020}{Timmermans et al. 2020}).
Ultimately, they act as cross-pollinators of co-creation methods,
approaches and perspectives between groups allowing to stimulate and
channel collective creativity so that new ideas constantly emerge
(\protect\hyperlink{ref-Jezierski2014}{Jezierski et al. 2014};
\protect\hyperlink{ref-Rayna2019}{Rayna and Striukova 2019}). Despite
how promising it is to pursue the innovation lab approach, it is
important to keep in mind that this is a response to keep up to
increasing and accumulating changes that we live today and where more
conventional approaches fall short
(\protect\hyperlink{ref-Zivkovic2018}{Zivkovic 2018};
\protect\hyperlink{ref-RezaeeVessal2021}{Vessal et al. 2021}). That is
to say, to embark on such an initiative implies to deal with
uncertainty, ambiguity and tensions that are inherent aspects of working
on such complex and changing conditions
(\protect\hyperlink{ref-Osorio2019}{Osorio, Dupont, Camargo, Palominos,
et al. 2019}; \protect\hyperlink{ref-Jezierski2014}{Jezierski et al.
2014}). This is why organizations willing to create their own ``lab''
need to be aware of the challenges and opportunities that this type of
initiative entails.

SI lab teams perform in permeable and changing environments where people
and organizations come and go depending on the problem that is being
addressed and the parties to whom it is relevant
(\protect\hyperlink{ref-Wascher2018}{Wascher et al. 2018}). These
ever-changing conditions demand for teams who value and practice
flexibility and agility in order to make the most of the ecosystem. Lab
teams should possess a wide range of competences that allow them to be
open to transitions and comfortable with ambiguities; use multiple
lenses to integrate multiple perspectives; and be able and willing to
work across-disciplines so resources can be mobilized in creative ways
(\protect\hyperlink{ref-Jezierski2014}{Jezierski et al. 2014};
\protect\hyperlink{ref-Puttick2014-Teams}{Puttick, Baeck, and Colligan
2014}; \protect\hyperlink{ref-Rayna2021}{Rayna and Striukova 2021}).
Still, literature seems to remain scarce when it comes to providing
guidance on which competences should SI lab teams have or focus their
development in order to succeed. In the next subsection, we explore this
issue in order to establish a common ground for what set of competences
a team for a SI lab should have.

\hypertarget{competence-and-social-innovation-labs}{%
\subsection{Competence and social innovation
labs}\label{competence-and-social-innovation-labs}}

From organizational and managerial perspectives, development of human
competences is a fundamental task in the path of innovation and
successful organizations
(\protect\hyperlink{ref-Leonard1995}{Leonard-Barton 1995}).
Understanding individual competences is key for enabling teams and
organizations to perform and adapt in rapidly changing conditions
(\protect\hyperlink{ref-Sandberg2000}{Sandberg 2000}). In general, the
concept of \emph{competence} is understood as the capability of an
individual to deliver sustainable and effective performance in a
specific domain, job, role, organizational context or situation
(\protect\hyperlink{ref-Mulder2014}{Mulder 2014}). A \emph{competence}
consists of various \emph{competencies} that coherently cluster a set of
knowledge, skills, attitudes and experience
(\protect\hyperlink{ref-Mulder2014}{Mulder 2014}). In that sense,
competence profiles are often used to represent the functional and
behavioral competencies that are required to successfully meet complex
demands in a particular context
(\protect\hyperlink{ref-Chatenier2010}{Chatenier et al. 2010}).

In the context of open innovation for instance,
(\protect\hyperlink{ref-Chatenier2010}{Chatenier et al. 2010}) proposed
in their work a competence profile for open innovation teams based on 20
semi-structured interviews and 2 focus groups with professionals that
had participated in open innovation projects in the agribusiness sector.
Based on their empirical findings, they built a profile composed of 4
main competence categories and 34 key competencies to accomplish three
main tasks of an open innovation team: (1) managing the
inter-organizational collaboration process, (2) managing the overall
innovation process and (3) creating new knowledge collaboratively. They
consider that a team having competence in self-management, interpersonal
management, project management and content management should be better
prepared to deal with the challenges behind those main tasks.

In a similar way, \protect\hyperlink{ref-Podmetina2015}{Podmetina et
al.} (\protect\hyperlink{ref-Podmetina2015}{2015}) proposed an open
innovation specialist profile based on a large-scale survey with 528
European companies. By inquiring on the required and desired
competencies that an employee should have for implementing open
innovation, they build a profile composed of six categories of
competencies: collaboration skills, interdisciplinary skills, methodic
skills, explorative skills, transformational skills and exploitative
skills. This work will subsequently lead to the proposition of a
competence model for open innovation in which direct links between
competencies, key activities and roles are made at the organizational
level (\protect\hyperlink{ref-Podmetina2018}{Podmetina et al. 2018}).
This holistic understanding of what constitutes a person's essentials
elements for performing in a determined task or role is instrumental for
assembling teams and training professionals
(\protect\hyperlink{ref-Mulder2014}{Mulder 2014}). Furthermore, a
competence profile can also be used as an assessment tool of ongoing
teams in order to spark reflecting processes
(\protect\hyperlink{ref-Sandberg2000}{Sandberg 2000}). This ultimately
allows managers for identifying whether there is room for improvement
and deciding what kind of actions are worth pursuing in order to enhance
a team's performance, especially in the complex and uncertain
circumstances such as the ones of facilitating innovation processes
(\protect\hyperlink{ref-Chatenier2010}{Chatenier et al. 2010}).
Nevertheless, although the studies conducted by
(\protect\hyperlink{ref-Chatenier2010}{Chatenier et al. 2010}) and
(\protect\hyperlink{ref-Podmetina2018}{Podmetina et al. 2018}) are
presented as specific but not unique to the open innovation context,
little has been studied in terms of SI and SI labs.

While the existing literature on SI labs constantly highlights the
importance of the lab team and the selection of the staff, most of
today's experiences and insights rely on generic statements such as the
need of people with mixed profiles and backgrounds to reflect the social
reality, with both traditional skills such as project management and
communications and innovation skills to get things done, or with
networking skills to gather participants and build connections
(\protect\hyperlink{ref-Puttick2014-Teams}{Puttick, Baeck, and Colligan
2014}; \protect\hyperlink{ref-Kieboom2015}{Kieboom, Exel, and Sigaloff
2015}; \protect\hyperlink{ref-Jezierski2014}{Jezierski et al. 2014}).
Acknowledging the importance of this issue,
\protect\hyperlink{ref-Wascher2018}{Wascher et al.}
(\protect\hyperlink{ref-Wascher2018}{2018}) gathered from the literature
a set of 14 competences which they proposed as key for a SI lab team
(see Table 1). They consider that the combination of all of these
competences should help the team to successfully manage and facilitate
cross-sector collaborations. Furthermore, these teams tend to be
relatively small, usually composed by a lab manager, administrative
staff and members dedicated to the lab-process facilitation
(\protect\hyperlink{ref-Wascher2018}{Wascher et al. 2018}). Yet, there
is no evidence that suggests what are the required roles for a SI lab
and further, which competences are needed to effectively perform those
roles.

\providecommand{\docline}[3]{\noalign{\global\setlength{\arrayrulewidth}{#1}}\arrayrulecolor[HTML]{#2}\cline{#3}}

\setlength{\tabcolsep}{2pt}

\renewcommand*{\arraystretch}{1.5}

\begin{longtable}[c]{|p{1.00in}|p{4.00in}}

\caption{SI Lab list of competences retrieved from [@Wascher2018]}\\

\hhline{~~}

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{CFCFCF}\raggedright}p{\dimexpr 1in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{\textbf{Competence}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{CFCFCF}\raggedright}p{\dimexpr 4in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{\textbf{Description}}}} \\



\endfirsthead

\hhline{~~}

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{CFCFCF}\raggedright}p{\dimexpr 1in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{\textbf{Competence}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{CFCFCF}\raggedright}p{\dimexpr 4in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{\textbf{Description}}}} \\

\endhead



\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 1in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{9}{9}\selectfont{\textcolor[HTML]{000000}{Project management}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 4in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{9}{9}\selectfont{\textcolor[HTML]{000000}{Competence in planning and implementing innovative projects; meeting legal requirements as well as financial expertise, contracts and agreements on the use of space}}} \\





\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 1in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{9}{9}\selectfont{\textcolor[HTML]{000000}{Moderation}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 4in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{9}{9}\selectfont{\textcolor[HTML]{000000}{Competence for integrating emerging ideas and orient projects}}} \\





\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 1in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{9}{9}\selectfont{\textcolor[HTML]{000000}{Mediation}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 4in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{9}{9}\selectfont{\textcolor[HTML]{000000}{Competence for helping project parties understand and focus on the important issues needed to reach a conflict resolution}}} \\





\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 1in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{9}{9}\selectfont{\textcolor[HTML]{000000}{Networking}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 4in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{9}{9}\selectfont{\textcolor[HTML]{000000}{Competence for building connections and relationships with local organizations}}} \\





\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 1in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{9}{9}\selectfont{\textcolor[HTML]{000000}{Participation}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 4in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{9}{9}\selectfont{\textcolor[HTML]{000000}{Competence in fostering mechanisms for the involvement of the parties in the project's decision-making processes}}} \\





\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 1in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{9}{9}\selectfont{\textcolor[HTML]{000000}{Communication}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 4in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{9}{9}\selectfont{\textcolor[HTML]{000000}{Competence for empathy, change of perspective and use of media in a clear, positive, conversational fashion}}} \\





\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 1in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{9}{9}\selectfont{\textcolor[HTML]{000000}{Self-organization}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 4in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{9}{9}\selectfont{\textcolor[HTML]{000000}{Competence for ambiguity and frustration tolerance, confidence and self-esteem}}} \\





\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 1in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{9}{9}\selectfont{\textcolor[HTML]{000000}{Intercultural}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 4in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{9}{9}\selectfont{\textcolor[HTML]{000000}{Competence in ensuring inclusivity throughout the project}}} \\





\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 1in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{9}{9}\selectfont{\textcolor[HTML]{000000}{Evaluation}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 4in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{9}{9}\selectfont{\textcolor[HTML]{000000}{Competence in the design of mechanisms for monitoring strategies and results}}} \\





\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 1in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{9}{9}\selectfont{\textcolor[HTML]{000000}{Research methods and interdisciplinary work}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 4in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{9}{9}\selectfont{\textcolor[HTML]{000000}{Competence for working under interdisciplinary environments using diverse research methods such as critical thinking, data analytics, social research, anthropology, etc.}}} \\





\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 1in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{9}{9}\selectfont{\textcolor[HTML]{000000}{Design methods and creative thinking}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 4in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{9}{9}\selectfont{\textcolor[HTML]{000000}{Competence in applying design methods such as design thinking, theory of change planning, etc.}}} \\





\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 1in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{9}{9}\selectfont{\textcolor[HTML]{000000}{Information and telecommunication techniques}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 4in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{9}{9}\selectfont{\textcolor[HTML]{000000}{Competence in technological techniques that provide support to the project development}}} \\





\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 1in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{9}{9}\selectfont{\textcolor[HTML]{000000}{Entrepreneurial thinking}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 4in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{9}{9}\selectfont{\textcolor[HTML]{000000}{Competence in project incubation processes and ventures}}} \\





\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 1in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{9}{9}\selectfont{\textcolor[HTML]{000000}{Systems thinking}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 4in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{9}{9}\selectfont{\textcolor[HTML]{000000}{Competence in addressing challenges in a holistic way and being able to examine the links and interactions between all the constituent elements}}} \\



\end{longtable}

\hypertarget{innovation-teams-roles}{%
\subsection{Innovation Teams \& Roles}\label{innovation-teams-roles}}

The idea of thinking on what are the roles or behaviors that are
required to facilitate the innovation process within an organization is
not new at all. One can refer to the notion of ``champion'' back in the
60's where the success of the innovation process was attributed to the
one single person who was willing to fail for a doubtful idea but
capable of reaching success (\protect\hyperlink{ref-Roberts1982}{Roberts
and Fusfeld 1982}; \protect\hyperlink{ref-Jenssen2004}{Jenssen and
Jørgensen 2004}). Nevertheless, the aim of reflecting on innovation
roles is no longer to create heroes that prevail against all odds.
Instead, it consists of building strong teams aware of their strengths
and weaknesses so that they can find ways to overcome the barriers in
the path of realizing the intended innovation process
(\protect\hyperlink{ref-Gemunden2007}{Gemünden, Salomo, and Hölzle
2007}). Indeed, innovation does not originate and sustain itself, but
rather through the people who make it happen through teams that push
their imagination, resilience and perseverance
(\protect\hyperlink{ref-Kelley2005}{Kelley and Littman 2005}).

It is in this sense that literature on innovation teams and roles has
evolved, as innovation processes have become more open, collaborative
and social, so it has been the case for the roles needed to facilitate
these processes. By diving into the literature of innovation teams we
intend to illustrate the diversity of roles that members of an
innovation team can have which subsequently could be of inspiration for
the set up of a lab team. Under this context, seven models of innovation
roles have been found in the literature (Figure 1) which will be now
discussed.

One of the earliest innovation role models that can be found is the one
proposed by (\protect\hyperlink{ref-Roberts1982}{Roberts and Fusfeld
1982}). Their proposition is composed of five roles needed to fulfil the
critical functions for a technology-based innovation process. These are
the \emph{idea generator}, the \emph{entrepreneur} or \emph{champion},
the \emph{project leader}, the \emph{gatekeeper} and the \emph{coach}.
The intention was to highlight those key functions that were not always
explicit in formal job structures. They also acknowledge that depending
on the size of the team or the organization, some roles need to be
fulfilled by more than one person, while some individuals can perform
more than one role, and that ultimately, the roles someone can fulfil
will change over a person's career. But beyond considering a role as
purely functional, it is even more important to ask how a person is
going to behave within a team. Under the premise that people's useful
behaviors can be grouped into a set of related clusters,
(\protect\hyperlink{ref-Belbin2010}{Belbin 2010}) condensed in her book
(originally published in 1993) the nine team roles which make an
effective contribution to team performance: \emph{plant, resource
investigator, coordinator, shaper, monitor evaluator, teamworker,
implementer, completer finisher and specialist}. Even though the Belbin
team roles are not exclusively for an innovation team, they represent an
important part of team theory that should be considered.

More recently, \protect\hyperlink{ref-Hering2005}{Hering and Phillips}
(\protect\hyperlink{ref-Hering2005}{2005}) presented eight innovation
roles making emphasis on those that are required for a generic
innovation process. They detailed the features of what can be expected
of these roles rather than just titles or job descriptions. According to
them, \emph{connector, librarian, framer, judge, prototyper, monitor}
and \_storyteller \_are the roles that should be sought to set-up an
innovation team. Organization's commitment and a belief system are also
considered critical for them in order to have the time and the resources
for innovation teams to deal with the uncertainty involved in any
innovation process. Alternatively,
\protect\hyperlink{ref-Kelley2005}{Kelley and Littman}
(\protect\hyperlink{ref-Kelley2005}{2005}) published their book
\emph{Ten Faces of Innovation} based on their experiences at IDEO. They
condensed ten persona descriptions as a way to inspire the roles that
members of an organization should play to foster creativity and
innovation. They consider that each role or persona helps to bring on
the table specific values, tools, skills and thus, it is important to
assure their presence in any innovation team. These ten roles are
grouped in \emph{learners} (anthropologist, experimenter and
cross-pollinator), \emph{organizers} (hurdler, collaborator and
director) and \emph{builders} (experience architect, set designer,
caregiver and storyteller).

Based on 104 interviews with representatives of German enterprises and
42 cases from questionnaires,
\protect\hyperlink{ref-Gemunden2007}{Gemünden, Salomo, and Hölzle}
(\protect\hyperlink{ref-Gemunden2007}{2007}) proposed a model to assess
whether the influence of certain innovation roles increase the success
of new product development under increasingly more open innovation
contexts. They pointed out that not only innovation and technological
experts are present (\emph{expert promoter} and \emph{process
promoter}), but strong leadership (\emph{project leader}) as well as
good external relationships (\emph{technology and market relationships
promoters}). Moreover, they emphasized the importance of having
institutional support in the form of \emph{power promoters}. In more
recent years, (\protect\hyperlink{ref-Goduscheit2014}{Goduscheit 2014})
builds on the work initiated by
(\protect\hyperlink{ref-Gemunden2007}{Gemünden, Salomo, and Hölzle
2007}). In this case, he seeks to further develop the concept of
innovation promoters. This notion is established on the basis that
innovation teams are meant to overcome the barriers and difficulties to
successful innovations. His interest was to explore the
inter-organizational dimension among the innovation roles proposed by
(\protect\hyperlink{ref-Gemunden2007}{Gemünden, Salomo, and Hölzle
2007}) by analyzing how they interact/perform with multiple
organizations. As a result, he further develops the innovation promoters
model by moving from the original six roles to a proposition of nine
roles: \emph{seniority, top-level representative, technological expert,
methodology expert, intra-organizational process, inter-organizational
process, project process, technology relationship and} market
relationship.

Finally, we refer to the very interesting work conducted by
(\protect\hyperlink{ref-Nystrom2014}{Nyström et al. 2014}). They also
build on the work realized by
(\protect\hyperlink{ref-Gemunden2007}{Gemünden, Salomo, and Hölzle
2007}) on analyzing the roles for an open innovation context but they
center their research on the influence these roles can have on
innovation networks. For this, they studied 26 living labs leading to a
final proposition of 17 roles that network actors can adopt or create
during an innovation project. The new roles identified are mostly
related to the users and the facilitators (\emph{e.g., co-creator,
orchestrator, contributor}), which correspond to living lab approaches
that encourage multi-stakeholder involvement. They also state the
importance of innovation roles to combine multiple perspectives due to
the increasing complexity of innovation projects. This is something that
relates to the more systemic and transdisciplinary approach that is
required on SI projects.

Throughout this literature review it is possible to observe that despite
the diversity of perspectives, processes or names, authors agree that
unbalanced teams and frequent changes can disturb how an innovation team
performs. This is a challenge that definitely should be considered in
the conformation of an innovation lab. However, none of the role models
establishes a direct link between the proposed roles and the adequate
competences that should allow a person to fulfil it. Nor any of the
identified studies is developed under the SI context. These elements are
taking into account the proposition of this article in the next section.

\hypertarget{methodology}{%
\section{Methodology}\label{methodology}}

Throughout the theoretical background presented before, the principles
behind the notion of SI labs have been explored, along with the dynamics
that lab teams are deemed to deal in such kind of context. Several
questions have been raised in terms on which roles would allow a SI lab
team to be better prepared to accomplish their mission. And further,
what set of competences would be necessary for these teams to thrive in
such conditions. Accordingly, seven innovation role models have been
retrieved from the literature on innovation teams as well as a set of 14
competences considered as key for SI lab teams. However, since research
on how managerial teams of innovation labs perform, and more
specifically those within the SI context remains unexplored, through
this study we intend to establish a connection between theory from
competences for innovation and innovation teams in order to hypothesize
on the essentials elements that should be considered in the process of
assembling a SI lab team. To this end, a four-stage process was designed
for conducting this research as shown in Figure 2.

Firstly, based on the literature review an adapted role model is
proposed. Since none of the previous role models for innovation teams
were rooted in SI nor innovation labs, we aim to take into consideration
the 14 identified competences to make a model proposition adapted to the
conditions of this research. Following this, the proposed model is
operationally defined as an assessment tool (online questionnaire) that
would ultimately be applied under a self-assessment approach. Given the
practical motivation behind this exploratory study, is to support 10
latin-american university teams to set up the lab team for their own SI
lab within the frame of the Erasmus+ Climate Labs project, we opted to
pursue a self-assessment approach with a two-fold purpose. Firstly, to
use the proposed approach to spark reflecting processes that would allow
the university teams to increase awareness of their current status
toward the expected roles. And secondly, to have a comprehensive role
characterization at the early stage of the project encompassing all the
lab teams according to their own perception of the degree of mastery of
such competences.

Due to the transcontinental nature of the Climate Labs project, the
application of the self-assessment approach was conducted virtually by
means of the online questionnaire and an online workshop. The proposed
approach was designed so each lab team member (including professors,
researchers, students, and administrative staff) could participate and
be part of the process. A total of 65 answers were received along with
the workshop results for each team. Results and insights are then
analyzed and discussed so conclusions can be made in order to provide
guidelines for the future of the Climate Labs project but also to the
further development of this study.

\hypertarget{proposition-of-a-competence-based-role-model-for-si-lab-teams}{%
\section{Proposition of a competence-based role model for SI Lab
teams}\label{proposition-of-a-competence-based-role-model-for-si-lab-teams}}

Despite several insights from empirical studies and different statements
of which functions or behaviours are possible to find in an innovation
team, the propositions and explanations fall short when it comes to the
specificity of innovation lab teams. We therefore believe that it is
reasonable to think that by establishing a clearer connection between
competences for SI labs and innovation team theory a model can be
proposed. First, drawing from the 14 competences proposed by
(\protect\hyperlink{ref-Wascher2018}{Wascher et al. 2018}), a
categorization was made thinking on which main functions could be
proposed. Based on the literature and according to knowledge and
experience of the authors, four categories of competences were
identified as illustrated in Table 2. This was done in terms of those
competences that contribute the most to one of the following functions:
(1) innovation process orchestration, (2) materialize systemic
solutions, (3) spark connections and new ideas, and those that
contribute to (4) organizing and measuring results.

\providecommand{\docline}[3]{\noalign{\global\setlength{\arrayrulewidth}{#1}}\arrayrulecolor[HTML]{#2}\cline{#3}}

\setlength{\tabcolsep}{2pt}

\renewcommand*{\arraystretch}{1.5}

\begin{longtable}[c]{|p{1.00in}|p{0.80in}|p{0.80in}|p{0.80in}|p{0.80in}}

\caption{Categorization of SI lab team competences}\\

\hhline{~~~~~}

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{CFCFCF}\raggedright}p{\dimexpr 1in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\textbf{Competence}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{CFCFCF}\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\textbf{Orchestrate Innovation Process}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{CFCFCF}\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\textbf{Materialize Systemic Solutions}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{CFCFCF}\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\textbf{Spark Connections \& Ideas}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{CFCFCF}\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\textbf{Organize and measure results}}}} \\



\endfirsthead

\hhline{~~~~~}

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{CFCFCF}\raggedright}p{\dimexpr 1in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\textbf{Competence}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{CFCFCF}\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\textbf{Orchestrate Innovation Process}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{CFCFCF}\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\textbf{Materialize Systemic Solutions}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{CFCFCF}\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\textbf{Spark Connections \& Ideas}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{CFCFCF}\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{\textbf{Organize and measure results}}}} \\

\endhead



\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 1in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{Project management}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{X}}} \\





\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 1in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{Moderation}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{X}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{}}} \\





\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 1in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{Mediation}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{X}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{}}} \\





\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 1in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{Networking}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{X}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{}}} \\





\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 1in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{Participation}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{X}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{}}} \\





\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 1in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{Communication}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{X}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{}}} \\





\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 1in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{Self-organization}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{X}}} \\





\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 1in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{Intercultural}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{X}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{}}} \\





\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 1in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{Evaluation}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{X}}} \\





\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 1in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{Research methods and interdisciplinary work}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{X}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{}}} \\





\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 1in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{Design methods and creative thinking}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{X}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{}}} \\





\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 1in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{Information and telecommunication techniques}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{X}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{}}} \\





\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 1in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{Entrepreneurial thinking}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{X}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{}}} \\





\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 1in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{Systems thinking}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{X}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.8in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{8}{8}\selectfont{\textcolor[HTML]{000000}{}}} \\



\end{longtable}

Then, in order to obtain a better sense of how these competences can be
better defined in terms of an innovation team and its roles, we compared
the four categories of competences with the innovation roles found in
the literature. At this point, we looked for similarities between the
roles and behaviors that have been historically recognized among
innovation teams with the competences for SI labs. We found that all
four categories are present in the seven innovation team role models
(see Table 3). It is worth remembering that our focus is this study lies
on the composition of the lab teams. This means that during this
comparison our attention remained on those roles at the core of the
management of the innovation lab. This is why a fifth category emerges
as ``external roles'' since they refer to users or external parties that
surround the lab network of stakeholders. We do not intend to diminish
the importance of these roles but on the contrary, we would like to
remark that indeed those external roles should definitely be considered
in depth in further stages of the research. However, considering the
relative small size that innovation lab teams usually have, especially
in early stages, we believe that these four categories constitute a
comprehensive and pragmatic basis for this study.

Based on the previous work, a competence-based role model for SI labs is
proposed (see Figure 3). This model is intended to establish a minimal
base of competences and roles that a SI lab team should bear in mind
during the implementation of their project. A general description for
each of the proposed roles is outlined below.

\hypertarget{facilitator-the-ones-who-set-the-tempo-and-orchestrate-peoples-collaboration}{%
\subsubsection{Facilitator -- The ones who set the tempo and orchestrate
people's
collaboration}\label{facilitator-the-ones-who-set-the-tempo-and-orchestrate-peoples-collaboration}}

This is probably the role that most distinguishes an innovation lab
team. The Facilitators are those who set the methodological tempo of the
projects within the lab. They just know what strategy, tool or dynamic
to use when divergence, convergence or a ``simple'' retrospective is
needed. They get people involved and seek collaboration in every
activity. When intentions and interests collide, facilitators keep
everyone focused on the greater purpose. No collaboration is possible
under discrimination and exclusion, so intercultural and
transdisciplinarity inclusiveness is a prime directive for them. What
could be expected from facilitators in a SI lab team:

\begin{itemize}
\tightlist
\item
  Design strategies, methods, and tools for orienting the innovation
  process through each project
  (\protect\hyperlink{ref-Belbin2010}{Belbin 2010};
  \protect\hyperlink{ref-Hering2005}{Hering and Phillips 2005};
  \protect\hyperlink{ref-Kelley2005}{Kelley and Littman 2005};
  \protect\hyperlink{ref-Gemunden2007}{Gemünden, Salomo, and Hölzle
  2007}; \protect\hyperlink{ref-Goduscheit2014}{Goduscheit 2014}).
\item
  Provide guidance and mentoring to stimulate professional and personal
  development of project teams and participants
  (\protect\hyperlink{ref-Robert1982}{\textbf{Robert1982?}};
  \protect\hyperlink{ref-Kelley2005}{Kelley and Littman 2005}).
\item
  Get people involved and encourage collaboration in every lab activity
  (\protect\hyperlink{ref-Kelley}{\textbf{Kelley?}};
  \protect\hyperlink{ref-Gemunden2007}{Gemünden, Salomo, and Hölzle
  2007}; \protect\hyperlink{ref-Nystrom2014}{Nyström et al. 2014};
  \protect\hyperlink{ref-Goduscheit2014}{Goduscheit 2014}).
\item
  Act as peacekeepers when conflict emerges, maintaining focus on
  project objectives and common goals
  (\protect\hyperlink{ref-Belbin2010}{Belbin 2010};
  \protect\hyperlink{ref-Hering2005}{Hering and Phillips 2005};
  \protect\hyperlink{ref-Nystrom2014}{Nyström et al. 2014}).
\item
  Ensure intercultural inclusiveness in all the lab projects or
  activities {[}\protect\hyperlink{ref-Kelley2005}{Kelley and Littman}
  (\protect\hyperlink{ref-Kelley2005}{2005}); Nystrom2014;
  Wascher2018{]}.
\end{itemize}

\hypertarget{maker-the-ones-who-make-things-happen}{%
\subsubsection{Maker -- The ones who make things
happen}\label{maker-the-ones-who-make-things-happen}}

Whomever comes with an idea usually needs support to make it real.
Makers gather a diverse set of skills that are instrumental to
understand every need and prototype any solution. They do not hesitate
in mixing disciplines or going to the field in order to understand the
problem. They shine at combining ideas and driving innovative concepts.
Makers are driven by experimentation, iterating through prototypes
crafted on available technologies. Acknowledging the complexity of each
problem in a systemic manner is part of the way they see the world. What
could be expected from makers in a SI lab team:

\begin{itemize}
\tightlist
\item
  Understand needs and problems through the combination of multiple
  research methods and settings
  {[}\protect\hyperlink{ref-Kelley2005}{Kelley and Littman}
  (\protect\hyperlink{ref-Kelley2005}{2005}); Nystrom2014;
  \protect\hyperlink{ref-Wascher2018}{Wascher et al.}
  (\protect\hyperlink{ref-Wascher2018}{2018}){]}.
\item
  Capture ideas, data, and any form of knowledge, restructure them and
  propose novel concept solutions
  (\protect\hyperlink{ref-Roberts1982}{Roberts and Fusfeld 1982};
  \protect\hyperlink{ref-Belbin2010}{Belbin 2010};
  \protect\hyperlink{ref-Hering2005}{Hering and Phillips 2005}).
\item
  Help to build physical (and digital) representations of solutions on
  an iterative basis (\protect\hyperlink{ref-Belbin2010}{Belbin 2010};
  \protect\hyperlink{ref-Hering2005}{Hering and Phillips 2005};
  \protect\hyperlink{ref-Kelley2005}{Kelley and Littman 2005};
  \protect\hyperlink{ref-Gemunden2007}{Gemünden, Salomo, and Hölzle
  2007}; \protect\hyperlink{ref-Nystrom2014}{Nyström et al. 2014};
  \protect\hyperlink{ref-Goduscheit2014}{Goduscheit 2014}).
\item
  Propose alternatives to address the complexity of every problem and
  the systemic impact of each solution
  (\protect\hyperlink{ref-Nystrom2014}{Nyström et al. 2014};
  \protect\hyperlink{ref-Wascher2018}{Wascher et al. 2018}). 3.
  Visionary -- The ones who push and connect
\end{itemize}

Refers to the kind of person who always finds an opportunity where no
one else sees it. These very same people also create networks easily.
They have an inherent talent to communicate their ideas while also
keeping themselves empathetic and sensitive to others. Their particular
entrepreneurial mindset keeps them moving forward looking to achieve
their vision of the future and beliefs. What could be expected from
visionaries in a SI lab team:

\begin{itemize}
\tightlist
\item
  Provide a constant flow of ideas and project opportunities for the lab
  (\protect\hyperlink{ref-Roberts1982}{Roberts and Fusfeld 1982};
  \protect\hyperlink{ref-Hering2005}{Hering and Phillips 2005};
  \protect\hyperlink{ref-Nystrom2014}{Nyström et al. 2014}).
\item
  Build connections with communities and stakeholders for establishing
  strong links between the lab, their stakeholders, and the territory
  (\protect\hyperlink{ref-Belbin2010}{Belbin 2010};
  \protect\hyperlink{ref-Hering2005}{Hering and Phillips 2005};
  \protect\hyperlink{ref-Kelley2005}{Kelley and Littman 2005};
  \protect\hyperlink{ref-Gemunden2007}{Gemünden, Salomo, and Hölzle
  2007}; \protect\hyperlink{ref-Nystrom2014}{Nyström et al. 2014};
  \protect\hyperlink{ref-Goduscheit2014}{Goduscheit 2014}).
\item
  Create an emotional connection with the people around the lab by
  communicating compelling stories from each project, event, success, or
  failure (\protect\hyperlink{ref-Hering2005}{Hering and Phillips 2005};
  \protect\hyperlink{ref-Kelley2005}{Kelley and Littman 2005};
  \protect\hyperlink{ref-Nystrom2014}{Nyström et al. 2014};
  \protect\hyperlink{ref-Goduscheit2014}{Goduscheit 2014}).
\item
  Go out and find the opportunities for the lab whether they are new
  alliances, funding options or showcase scenarios
  (\protect\hyperlink{ref-Belbin2010}{Belbin 2010};
  \protect\hyperlink{ref-Gemunden2007}{Gemünden, Salomo, and Hölzle
  2007}; \protect\hyperlink{ref-Goduscheit2014}{Goduscheit 2014};
  \protect\hyperlink{ref-Wascher2018}{Wascher et al. 2018}).
\end{itemize}

\hypertarget{manager-the-ones-who-keep-everything-on-track}{%
\subsubsection{Manager -- The ones who keep everything on
track}\label{manager-the-ones-who-keep-everything-on-track}}

Innovation lab initiatives can often be seen as fuzzy and accessory.
Managers are meant to organize, give sense, and value everything that
happens in a lab. Used to dealing with uncertain conditions, lab
managers are exceptionally adept at handling innovation projects. They
keep track of everything, detecting patterns of what works and what does
not. Their ability to watch metrics and results allows them to build
bridges with managerial counterparts from other institutions to defend
the role of the lab in their ecosystem. Lab managers are capable of
decision-making, which is key in the highly complex contexts where the
lab normally operates. What could be expected from managers in a SI lab
team:

\begin{itemize}
\tightlist
\item
  Handle technical, financial, and legal issues of the lab and its
  projects (\protect\hyperlink{ref-Gemunden2007}{Gemünden, Salomo, and
  Hölzle 2007}; \protect\hyperlink{ref-Nystrom2014}{Nyström et al.
  2014}; \protect\hyperlink{ref-Goduscheit2014}{Goduscheit 2014}).
\item
  Contribute to project planning keeping the balance between visionary
  solutions and achievable goals
  (\protect\hyperlink{ref-Roberts1982}{Roberts and Fusfeld 1982};
  \protect\hyperlink{ref-Belbin2010}{Belbin 2010};
  \protect\hyperlink{ref-Kelley2005}{Kelley and Littman 2005};
  \protect\hyperlink{ref-Nystrom2014}{Nyström et al. 2014}).
\item
  Implement monitoring and assessment mechanisms to track the lab's
  evolution and communicate results
  {[}\protect\hyperlink{ref-Roberts1982}{Roberts and Fusfeld}
  (\protect\hyperlink{ref-Roberts1982}{1982});
  \protect\hyperlink{ref-Belbin2010}{Belbin}
  (\protect\hyperlink{ref-Belbin2010}{2010}); Hering2005{]}.
\item
  Have a strong belief in themselves which helps them make decisions for
  the sustainability of the lab and its ecosystem
  (\protect\hyperlink{ref-Belbin2010}{Belbin 2010};
  \protect\hyperlink{ref-Gemunden2007}{Gemünden, Salomo, and Hölzle
  2007}; \protect\hyperlink{ref-Wascher2018}{Wascher et al. 2018}).
\end{itemize}

\hypertarget{design-of-the-instrument-a-self-assessment-approach}{%
\subsubsection{Design of the instrument: A self-assessment
approach}\label{design-of-the-instrument-a-self-assessment-approach}}

The next step was to operationalize the proposed role model. Rather than
relying entirely on a rationalistic approach where functional roles are
predefined and team members are expected to fit into those assumptions,
we seek to provide a reflection tool for lab teams to raise awareness as
they design and implement their SI lab
(\protect\hyperlink{ref-Sandberg2000}{Sandberg 2000}). In this sense, a
self-assessment approach is then instrumental to understand the degree
in which a person considers him/herself competent in a determined task
or situation while at the same time reflecting processes are sparked
(\protect\hyperlink{ref-Sandberg2000}{Sandberg 2000};
\protect\hyperlink{ref-Allen2005}{Allen and Velden 2005}). An important
advantage of following a self-assessment approach lies in the fact that
individuals have access to information about themselves that might
otherwise be more difficult to access
(\protect\hyperlink{ref-Allen2005}{Allen and Velden 2005}). However, it
is also important to be aware that at the same time this reliance on
self-knowledge often leads to problems in reliability and validity
results (\protect\hyperlink{ref-Ward2000}{\textbf{Ward2000?}}).

For this reason, the operationalization of our theoretical model
included firstly, the definition of short descriptions for each
competence (see Appendix A) which constitute the base for an online
questionnaire as the first step of the self-assessment. Accordingly, the
online questionnaire has 23 questions directly related to the model
criteria plus 6 general questions for profiling purposes and one open
question at the end for collecting feedback for a total of 30.
Furthermore, based on this structure, the questionnaire was also made
available in the three languages, namely Spanish, Portuguese and
English. In terms of the self-assessment levels, a numeric scale from 0
to 5 with anchoring phrases was applied (see Table 4). Being 1 the level
where respondents understand what the competence is but they do not
practice it, and 5, where they believe they have mastered the competence
and are capable of developing new ways of applying it
(\protect\hyperlink{ref-Dreyfus1986}{Dreyfus and Dreyfus 1986}). The 0
level was introduced for those who are not aware or do not understand
what the competence is about, allowing to reduce unintentional
measurement errors (\protect\hyperlink{ref-Allen2005}{Allen and Velden
2005}). In order to deal with ambiguity or lack of clarity, a test was
done with several members of the consortium prior to applying the study
with the latin-american teams. Adjustments were then made based on the
feedback and test results. The english version of the final
questionnaire can be accessed at
\url{https://forms.gle/iZ1Lwyt1KUqrerb57}.

Secondly, a retrospective workshop was designed as a post
self-assessment intervention in order to ask participants to reflect as
a team on the results of the questionnaire. One of the main hypotheses
considers that a competence profile resulting from the self-assessment
enables a discussion among lab team members concerning: a) the strengths
and weaknesses they may have as a team; b) the potential role(s) that
each member could play in the lab; c) the aspects to develop and
possible strategies for doing so. In this way, lab teams will be
stimulated to reflect on their present conception of their lab and what
needs to be strengthened (\protect\hyperlink{ref-Sandberg2000}{Sandberg
2000}).

\hypertarget{applying-the-self-assessment-approach}{%
\section{Applying the self-assessment
approach}\label{applying-the-self-assessment-approach}}

\hypertarget{sample-and-data-collection}{%
\subsection{Sample and data
collection}\label{sample-and-data-collection}}

As previously mentioned, this study is conducted as part of the Climate
Labs project, an Erasmus+ initiative whose aim is to strengthen the
applied research and innovation capacities of a group of latinamerican
universities via the implementation of SI labs for mitigation and
adaptation to Climate Change. One of the project main goals is to build
multidisciplinary teams within universities that will lead the design of
their Climate Lab. Therefore the self-assessment was conceived as a
starting point in this project. The methodology was applied with 10
latinamerican universities from Colombia (5), Mexico (2) and Brazil (3)
with the process lasting between the months of April and August 2020.

Firstly, members of each team were asked to self-assess their SI lab
competences individually by means of the online questionnaire during
April and May. Later in June, a webinar was held with all teams in order
to share the preliminary results and introduce the retrospective
activity as the second part of the methodology. Each university team was
handed a customized report composed mainly by their competence profile
according to our four generic role propositions: Visionary, Maker,
Facilitator and Manager. The intention was to trigger a reflecting
process so strengths and improvement points could be identified and
discussed among the team. Based on the results, they were also asked to
state to which of the proposed roles they identified the most and the
least. And finally, they were also requested to ideate on potential
strategies for filling their gaps as a lab team.

A total of 65 participants completed both the questionnaire and the
workshop. Lab Team size oscillates between 4 and 13 members. Among
respondents there are faculty members, administration staff, researchers
and students. The latters being only represented by three undergraduate
students and one doctoral student. This was mainly because of
difficulties imposed by the global health situation due to COVID-19,
with repercussions on the respective academic calendars of the
universities. Nevertheless, the sample is still representative of the
people that will lead the lab implementation which guarantees that
insights from this analysis will offer a real picture of the early state
of a SI lab team. Table 5 summarizes the team composition per
university.

\hypertarget{instrument-internal-consistency}{%
\subsection{Instrument internal
consistency}\label{instrument-internal-consistency}}

It is of interest to have a measure of the instrument's consistency.
Then, we chose to calculate Cronbach's alpha coefficient values since
they have been widely used for construct validation purposes
(\protect\hyperlink{ref-Taber2018}{Taber 2018}). Due to the
multidimensional nature of the questionnaire, alpha coefficient was
calculated for the groups of items associated with each role. Alpha for
maker and manager are lower, 0.773 and 0.787 respectively, but still
over 0.7 which is suggestive for consistency. Also alpha for visionary
has an acceptable coefficient of 0.824 whereas for facilitator is higher
with 0.906. Far from reaching any conclusion in terms of reliability,
results only indicate an acceptable level of interrelatedness (or
equivalence) among the grouped competences for each role, something we
consider positive at this exploratory stage and the subsequent steps of
this research.

\hypertarget{competence-self-assessment-results}{%
\subsection{Competence self-assessment
results}\label{competence-self-assessment-results}}

Figure 4 presents the self-assessment results for the whole sample
(black line). Three additional profiles are shown (green, blue and
orange dotted lines) which correspond to the three cases of lab teams
that will be analyzed further in this section. Here, the radar is used
to visualize the median-based profiles as a way to provide a measure of
the central tendency among the sample (or specific team). The median is
shown for each one of the 23 questions to help the teams to infer where
strengths or weaknesses might be. From a general perspective, there are
some points worth describing. The overall sample of participants
declared to have a high level of competence being able to guide others
(level 4) in moderation, research methods, networking and
self-organization. This could be expected since there is a significant
presence of professors and directors among the respondents which could
have already more training and experience than team members in other
positions. However, one result worth highlighting is the low level of
competence in ICT and entrepreneurial thinking, suggesting the existence
of common gaps that could pose challenges for the future development of
a SI lab.

\hypertarget{retrospective-results-a-cross-case-analysis}{%
\subsection{Retrospective results: A cross-case
analysis}\label{retrospective-results-a-cross-case-analysis}}

Based on the self-assessment results lab teams were invited to reflect
on their competence profile while the role model was introduced along
with the workshop steps. Due to space limitations, results from the
second part are described through three cases, namely, Labs H, I and J.
These cases were chosen because as a result of the self-assessment, they
stated that they lacked or had a low representation of specific roles
among their team members. This allows us to frame the analysis in terms
of what kind of conceptions the methodology elicits when specific roles
are missing. A cross-case analysis is then conducted to discuss
noteworthy observations and common elements in how lab teams interpreted
results, self-positioned and what kind of actions or improvement points
were identified. Table 6 presents a summary of the retrospective results
for the three chosen cases. Our attention turned into analyzing and
understanding how participants related to our role model and which
self-declared roles emerged for each case. We were also interested in
exploring how consistent (or not) are the competence profile results
with the self-declared roles along with the identified improvement
actions. For instance, do teams missing a specific role propose
corresponding actions to fill that gap? Besides engaging or involving
new people to the team, what other kind of reflections or improvement
strategies the methodology elicits? With these questions in mind, the
remaining of this section discusses some of the main insights.

The first part of the activity consisted in declaring to which role each
participant related the most according to the role model and their
competence profile. Some participants declared to be identified with
multiple roles whereas others with a single one. Among the selected
cases for this analysis, each one of them has a relative low
representation for a certain role. For instance, Lab I do not have any
self-declared maker. Lab J has only two visionaries and Lab H has only
one facilitator and one maker. It is interesting to see whether
strengths or weaknesses correspond or not to their team profile. This is
important since having no specific person identified or related to an
specific role does not necessarily mean that they are not competent as a
team in that regard.

In terms of strengths, the exercise seems to elicit both direct links
from their team profile where high levels of competence are identified
but also enable them to share information regarding additional key
points that characterize them as a team. For example, aspects such as
how long have they been working as a team, the degree of
institutionalization of SI and entrepreneurship at the university and
their strong links with the state government and local actors were
highlighted by Lab I as their main strengths. Perhaps one remarkable
characteristic is the fact that, as they stated, ``all our students have
a mandatory federal duty to dedicate 480 hours for social community
service, where we can count on their talent and work.'' This is
something that could prove useful for the lab once it is operating. The
team of Lab J acknowledges themselves as ``a Maker Team with great
research capacity'' as clearly the majority of them self-identified as
makers which also corresponds to high levels of competence in research
and design according to their global profile (see Figure 4). They also
emphasize the ``good positioning'' of several of their members in terms
of management and decision-making at their university. Remarkably,
despite their team profile evidencing a high level of competence for
facilitation (see Figure 4), only one member of Lab H self-identified as
facilitator, but still, among their strengths they consider themselves
``inclusive and respectful in moderating differences.'' They also
outline as strong points systemic thinking, project management,
pedagogical methods, networking and communications as well as their
knowledge in environmental disciplines and sustainable development.

As already perceived in the global profile, competence in ICT is a
common weakness that also appears in the retrospective for all three
cases. In addition, Lab H and J identify entrepreneurial thinking as
limited or with low experience whereas Lab I and also H raised concerns
on the slow and complicated administrative processes at the university.
More in particular, according to their competence profile, team members
of Lab H acknowledge a lack of experience in designing and prototyping
in innovation projects, and further, in building solutions to real
problems in multi stakeholder contexts. Similarly, for the Lab I team,
this is their first experience with a lab that combines SI and climate
change. They consider this a weakness as SI practices are not yet
commonly included in the academic work of other faculties. In addition
they state the lack of an emergency culture regarding climate action
could be a barrier when it comes to bringing changes or actions in their
institution at an operational level. Lastly, the team of Lab J expressed
the need of consolidating their visionary capabilities along with
systemic thinking and evaluation methodologies. They also pointed out
time management and the limited number of people available as weaknesses
since team members are also engaged in several additional projects at
the same time.

Moving forward in the retrospective, teams were then asked to synthesize
what competences they need to develop toward the design and
implementation of their SI lab. Besides reaching some direct conclusions
in terms of specific missing competences such as ICT or entrepreneurial
thinking, the activity also enabled them to elicit some less evident
conclusions. For instance, team members of Lab I stated that they are
missing maker roles with strengths in ICT, creative thinking and systems
thinking, as well as they need to develop a participation culture.
However, they also expressed as one of their conclusions ``we already
have the strengths of the other roles but we must keep increasing our
knowledge and expand onto other roles.'' They also agreed that having a
vision or a bigger picture of the scope of their SI lab is a key step
they need to develop before moving forward. On top of that, they
recognize it is important to also develop adapted project management
competence which recognizes and urges the nature of a SI project. This
goes in line also with members of Lab J who concluded that despite
having a high representation of the manager role, they still need a
project manager due to their time management issues. Teams also
identified competences or roles to develop according to their thematic
focus. Lab J team estimates the need to develop knowledge in Green
Economy strategies while Lab H highlights the importance for developing
competence in living lab and SI for climate change.

The final step in the retrospective was oriented toward the
identification of strategies and actions to improve or overcome the
elicited weaknesses or barriers. All three cases raised the importance
of involving students in this process where research group seedbeds,
social service or internships are identified as possible mechanisms to
this end. A second common strategy is to develop missing or weak
competences through training and learning communities. Team members seem
to recognize the experimental and learning nature of their SI lab
project thus potentially allowing for the consolidation of both
individual and team competences. More particularly, Lab H and J outlined
collaboration with other universities and actors in their ecosystem as a
source for fostering knowledge exchange and advice. Whereas Lab I
emphasized the importance of creating an identity around the lab so
members and non members could feel included and invited to be part of
it.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

The purpose of this study is to propose a methodological approach for
guiding the creation of SI lab teams. Since (social) innovation labs are
still an emergent field in the academic literature, this article
contributes to the literature by connecting the theory of innovation
teams with the competences for the management of SI processes. In this
way, a competence-based role model was proposed focusing on four main
roles to be taken into account as a starting point for SI lab
initiative. Although the literature recognizes the multiplicity of roles
in innovation teams and networks, in this study we chose to focus
strictly on the management point of view in the early stages of an
innovation lab. Prior studies have highlighted that facilitating
collaborative processes across disciplines, dealing with ambiguities,
integrating multiple perspectives, and building and sustaining a shared
intent in multi-stakeholder projects are some of the major challenges
that innovation lab teams confront today. Therefore, by focusing on the
facilitation, making, visionary and managerial facets of lab management,
innovation lab teams can find in this paper a deeper understanding of
the roles and behaviors that could effectively assist in the
implementation of the intended innovation processes.

Following an interpretative approach, the model was operationalized. As
a result, the proposed methodology consists of a self-assessment tool
coupled with a retrospective workshop. To the extent that the roles
required for an innovation lab have not yet been defined in the
literature, the proposed model was considered as the basis on which
teams can not only position themselves at the individual level in terms
of their competences and their relation to the proposed roles, but also
at the team level in order to identify and propose improvement actions.
Despite their relative small size, volatiness and sometimes informal
conditions, past experiences have shown that innovation lab initiatives
need to develop a human resource strategy since it becomes fundamental
when it comes to bringing the people with the right mindset to deal with
social challenges (\protect\hyperlink{ref-Osorio2020}{Osorio et al.
2020}; \protect\hyperlink{ref-Timeus2018}{Timeus and Gascó 2018}).
Ultimately, this work calls for lab teams to not only focus their
efforts on the projects and the different parties they interact with but
also to develop a ``reflectum practicum'' dynamic for sparking team
development and cohesion (\protect\hyperlink{ref-Sandberg2000}{Sandberg
2000}). In this sense, our methodological approach in application should
be seen not only as a tool for competence development but also for
closing the intent-gap within a team
(\protect\hyperlink{ref-Gratton1994}{Gratton 1994};
\protect\hyperlink{ref-Mantere2006}{\textbf{Mantere2006?}}), something
key for SI lab initiatives to thrive
(\protect\hyperlink{ref-Rayna2019}{Rayna and Striukova 2019};
\protect\hyperlink{ref-Osorio2019}{Osorio, Dupont, Camargo, Palominos,
et al. 2019}).

On top of that, it is important to bear in mind the idiosyncratic nature
of innovation labs which is why our approach can be also used by
researchers and practitioners to elicit conceptions in which lab teams
project the competences required for their lab according to their own
experience and context. SI labs continue to proliferate around the
world, but particularly the university context seems to find in this
type of initiative a means to stimulate deep transformations and enhance
its role as an agent of change in stressful times
(\protect\hyperlink{ref-Camargo2021}{Camargo, Morel, and Lhoste 2021};
\protect\hyperlink{ref-Petersen2021}{Petersen and Kruss 2021}). This is
precisely the case of the Climate Lab project and the cases that were
analyzed in this study. In application, the self-assessment helps to
have a ``picture'' of the lab teams where specific strengths and
weaknesses of this type of teams are revealed, but which are also
associated with the intrinsic conditions of the universities to which
they belong. In this way, this tool contributes to mapping aspects where
to focus attention to ensure the future development of a
university-hosted innovation lab, but also to identify and prioritize
the areas where it can generate greater value. Similarly, the
recognition and socialization of competence gaps seems to evoke the
identification of opportunities for collaboration with other labs and
institutions. This could eventually lead to inter-collaboration between
innovation labs by complementary capabilities and/or services
(\protect\hyperlink{ref-Memon2018}{Memon et al. 2018}).

Notwithstanding the implications of the self-assessment approach in SI
labs research and practice, this exploratory work comes with several
limitations that should be addressed in future research. Although the
competence-based role model proposed in this paper considers a broad
spectrum of the literature on roles and competences for innovation,
there is still plenty of room to investigate what are the underlying
competencies for each role and each competence, in terms of knowledge,
skills, aptitudes and experience
(\protect\hyperlink{ref-Chatenier2010}{Chatenier et al. 2010}). Further
theoretical and empirical research could explore the dynamics involved
in the management of socially innovative processes and how they differ
from open innovation teams, for example
(\protect\hyperlink{ref-Podmetina2018}{Podmetina et al. 2018}). In
addition, any self-assessment approach suffers from subjectivity.
Therefore, the results of the questionnaire should not be seen as
absolute statements, but on the contrary, they should be interpreted by
providing the appropriate space for collective reflection at the lab
team level. The use of a competence profile should motivate the
development of human talent, a process that is composed of chained
events rather than one single intervention at the start of an innovation
lab project (\protect\hyperlink{ref-Sandberg2000}{Sandberg 2000}).

Lastly, our choice of only four roles may be perceived as excluding due
to the multiplicity of behavioral and functional roles recognized in the
field of innovation. However, this work does not seek to overlook the
importance of previous studies, but on the contrary to open new research
paths to deepen the analysis from the management perspective of an
innovation lab and the talent in charge of fulfilling its mission. In
practical terms, this work offers a starting point for nascent lab teams
with limited resources, which is often the case for SI labs,
particularly in contexts such as Latin America
(\protect\hyperlink{ref-Osorio2020}{Osorio et al. 2020}). Future
research efforts should be focused on the evolution of role profiles
over time and which competences prevail in projects addressing social
challenges (\protect\hyperlink{ref-Rayna2021}{Rayna and Striukova 2021})
(i.e.~what does it mean to be a facilitator, maker, visionary or manager
for climate change adaptation and mitigation?). Further empirical
studies could also assess which aspects of the proposed model are most
influential in the performance of an innovation lab
(\protect\hyperlink{ref-Caccamo2020}{Caccamo 2020}).

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

This article explores the competences and team roles that could guide
the conformation of a SI lab team. Consequently, it develops a
competence-based role model composed of an existing set of 14
competences for SI gathered from the literature, which are then
organized under 4 generic roles for an innovation lab team. The
methodology, composed by an online self-assessment questionnaire plus a
retrospective activity, provides an interpretative approach for SI lab
teams to position themselves in terms of their competence profile and
the proposed roles while improvement strategies are identified. In
application with 10 lab teams from latinamerican universities, the
methodology proves instrumental for eliciting a team's strengths and
weaknesses but also key actions that could guide team conformation.
Results not only contribute to build knowledge from the team and
management point of view on the (emergent) SI lab literature but also
share detailed conceptions and experiences from nascent lab teams in the
latinamerican context.

The study does not come without limitations. The self-assessment
approach is often criticized due to the subjectivity of their results,
something that could be problematic looking for reliability and results
validation. However, the purpose of this work does not rely on producing
generalizable results on which roles a SI lab should have and its
underlying competences, but rather to provide a methodological approach
to support the assembling of such heterogeneous teams. Finally, further
research efforts remain in exploring in greater depth how team roles
perform and interact in an innovation lab so functional and behavioral
roles can be better understood. Furthermore, adding a processual
perspective to this study could provide insights on changes in roles
along with an innovation lab's evolution.

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

This work is part of the Climate Labs Project co-financed by the
European Commission through the Erasmus+ Programme under Grant Agreement
No.~610032. We would like to thank all members of the Climate Labs
consortium for their openness and active participation in the proposed
activities that make this study possible. This study also received
financial support from COLCIENCIAS Grant Program 785-2017 which funds
Ferney Osorio's PhD thesis research.

\newpage

\hypertarget{bibliography}{%
\section*{Bibliography}\label{bibliography}}
\addcontentsline{toc}{section}{Bibliography}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-Allen2005}{}%
Allen, J. P., and R. K. W. Van Der Velden. 2005. {``The Role of
Self-Assessment in Measuring Skills,''} January, 1--25.
\url{https://api.semanticscholar.org/CorpusID:153922219\%20https://cris.maastrichtuniversity.nl/en/publications/the-role-of-self-assessment-in-measuring-skills}.

\leavevmode\hypertarget{ref-Belbin2010}{}%
Belbin, R. Meredith. 2010. \emph{Team Roles at Work, Second Edition}.
\emph{Team Roles at Work, Second Edition}. Second. Taylor; Francis.
\url{https://doi.org/10.4324/9780080963242}.

\leavevmode\hypertarget{ref-Bjorklund2017}{}%
Björklund, Tua, Miko Laakso, Senni Kirjavainen, and Kalevi Ekman. 2017.
\emph{Passion-Based Co-Creation}. Aalto University.
\url{https://aaltodoc.aalto.fi/handle/123456789/29068}.

\leavevmode\hypertarget{ref-Bloom2016}{}%
Bloom, Louise, and Romily Faulkner. 2016. {``Innovation Spaces: Lessons
from the United Nations.''} \emph{Third World Quarterly} 37 (August):
1371--87. \url{https://doi.org/10.1080/01436597.2015.1135730}.

\leavevmode\hypertarget{ref-Caccamo2020}{}%
Caccamo, Marta. 2020. {``Leveraging Innovation Spaces to Foster
Collaborative Innovation.''} \emph{Creativity and Innovation
Management}, January, caim.12357.
\url{https://doi.org/10.1111/caim.12357}.

\leavevmode\hypertarget{ref-Camargo2021}{}%
Camargo, Mauricio, Laure Morel, and Pascal Lhoste. 2021. {``Progressive
University Technology Transfer of Innovation Capabilities to SMEs: An
Active and Modular Educational Partnership.''} In \emph{New Perspectives
in Technology Transfer. FGF Studies in Small Business and
Entrepreneurship.}, edited by Dana Mietzner and Christian Schultz, 1st
ed., 181--205. Springer, Cham.
\url{https://doi.org/10.1007/978-3-030-61477-5_11}.

\leavevmode\hypertarget{ref-Chatenier2010}{}%
Chatenier, Elise du, Jos A. A. M. Verstegen, Harm J. A. Biemans, Martin
Mulder, and Onno S. W. F. Omta. 2010. {``Identification of Competencies
for Professionals in Open Innovation Teams.''} \emph{R and D Management}
40 (June): 271--80.
\url{https://doi.org/10.1111/j.1467-9310.2010.00590.x}.

\leavevmode\hypertarget{ref-DeCusatis2008}{}%
DeCusatis, Casimer. 2008. {``Creating, Growing and Sustaining Efficient
Innovation Teams.''} \emph{Creativity and Innovation Management} 17
(June): 155--64. \url{https://doi.org/10.1111/j.1467-8691.2008.00478.x}.

\leavevmode\hypertarget{ref-Delgado2020}{}%
Delgado, Lorena, Daniel Galvez, Alaa Hassan, Pedro Palominos, and Laure
Morel. 2020. {``Innovation Spaces in Universities: Support for
Collaborative Learning.''} \emph{Journal of Innovation Economics \&
Management} 31: 123--53. \url{https://doi.org/10.3917/jie.pr1.0064}.

\leavevmode\hypertarget{ref-Dias2019}{}%
Dias, Joana, and Maria Partidário. 2019. {``Mind the Gap: The Potential
Transformative Capacity of Social Innovation.''} \emph{Sustainability
(Switzerland)} 11: 1--17. \url{https://doi.org/10.3390/su11164465}.

\leavevmode\hypertarget{ref-Dreyfus1986}{}%
Dreyfus, Hubert L., and Hubert L. Dreyfus. 1986. {``Five Steps from
Novice to Expert.''} In \emph{Mind Over Machine: The Power of Human
Intuition and Expertise in the Era of the Computer}, 231. Free Press.

\leavevmode\hypertarget{ref-Gemunden2007}{}%
Gemünden, Hans Georg, Sören Salomo, and Katharina Hölzle. 2007. {``Role
Models for Radical Innovations in Times of Open Innovation.''}
\emph{Creativity and Innovation Management} 16 (December): 408--21.
\url{https://doi.org/10.1111/j.1467-8691.2007.00451.x}.

\leavevmode\hypertarget{ref-Goduscheit2014}{}%
Goduscheit, René Chester. 2014. {``Innovation Promoters - a Multiple
Case Study.''} \emph{Industrial Marketing Management} 43 (April):
525--34. \url{https://doi.org/10.1016/j.indmarman.2013.12.020}.

\leavevmode\hypertarget{ref-Gratton1994}{}%
Gratton, Lynda. 1994. {``Implementing Strategic Intent: Human Resource
Processes as a Force for Change.''} \emph{Business Strategy Review} 5
(March): 47--66.
\url{https://doi.org/10.1111/j.1467-8616.1994.tb00069.x}.

\leavevmode\hypertarget{ref-Gregoire2016}{}%
Gregoire, Maud. 2016. {``Exploring Various Approaches of Social
Innovation: A Francophone Literature Review and a Proposal of Innovation
Typology.''} \emph{RAM. Revista de Administração Mackenzie} 17
(December): 45--71.
\url{https://doi.org/10.1590/1678-69712016/administracao.v17n6p45-71}.

\leavevmode\hypertarget{ref-Gryszkiewicz2016}{}%
Gryszkiewicz, Lidia, Ioanna Lykourentzou, and Tuukka Toivonen. 2016.
{``Innovation Labs: Leveraging Openness for Radical Innovation?''}
\emph{Journal of Innovation Management} 4 (March): 68--97.
\url{https://papers.ssrn.com/sol3/papers.cfm?abstract_id=2556692}.

\leavevmode\hypertarget{ref-Hellstrom2002}{}%
Hellström, Tomas, Merle Jacob, and Ulf Malmquist. 2002. {``Guiding
Innovation Socially and Cognitively: The Innovation Team Model at
Skanova Networks.''} \emph{European Journal of Innovation Management} 5
(September): 172--80. \url{https://doi.org/10.1108/14601060210436745}.

\leavevmode\hypertarget{ref-Hering2005}{}%
Hering, Dean, and Jeffrey Phillips. 2005. {``Innovation Roles the People
You Need for Successful Innovation.''} NetCentrics Corporation.

\leavevmode\hypertarget{ref-Jenssen2004}{}%
Jenssen, Jan Inge, and Geir Jørgensen. 2004. {``How Do Corporate
Champions Promote Innovations?''} \emph{International Journal of
Innovation Management} 08 (March): 63--86.
\url{https://doi.org/10.1142/s1363919604000964}.

\leavevmode\hypertarget{ref-Jezierski2014}{}%
Jezierski, E, J Harvey, L Hansen, M Takeuchi, R Sinha, K M M Kieboom,
and D Edwards. 2014. \emph{Labcraft: How Social Labs Cultivate Change
Through Innovation and Collaboration}. Labcraft Publishing.
\url{https://www.kl.nl/en/publications/labcraft-innovation-labs-cultivate-change/}.

\leavevmode\hypertarget{ref-Kelley2005}{}%
Kelley, Tom, and Jonathan Littman. 2005. \emph{The Ten Faces of
Innovation : IDEO's Strategies for Beating the Devil's Advocate \&
Driving Creativity Throughout Your Organization}. Currency/Doubleday.

\leavevmode\hypertarget{ref-Kieboom2015}{}%
Kieboom, Marlieke, Thijs van Exel, and Chris Sigaloff. 2015. {``Lab
Practice: Creating Spaces for Social Change.''} Kennisland.
\url{https://www.kl.nl/en/publications/lab-practice-creating-spaces-for-social-change/}.

\leavevmode\hypertarget{ref-Kratzer2004}{}%
Kratzer, Jan, Oger Th A. J. Leenders, and Jo M. L.van Engelen. 2004.
{``Stimulating the Potential: Creative Performance and Communication in
Innovation Teams.''} \emph{Creativity and Innovation Management} 13
(March): 63--71. \url{https://doi.org/10.1111/j.1467-8691.2004.00294.x}.

\leavevmode\hypertarget{ref-Kratzer2006-A}{}%
Kratzer, Jan, Roger Th.A. J. Leenders, and Jo M. L. Van Engelen. 2006.
{``Managing Creative Team Performance in Virtual Environments: An
Empirical Study in 44 r\&d Teams.''} \emph{Technovation} 26 (January):
42--49. \url{https://doi.org/10.1016/j.technovation.2004.07.016}.

\leavevmode\hypertarget{ref-Kratzer2006-B}{}%
Kratzer, Jan, Roger Th.A. J. Leenders, and Jo M. L. van Engelen. 2006.
{``Team Polarity and Creative Performance in Innovation Teams.''}
\emph{Creativity and Innovation Management} 15 (March): 96--104.
\url{https://doi.org/10.1111/j.1467-8691.2006.00372.x}.

\leavevmode\hypertarget{ref-Lake2016}{}%
Lake, Danielle, Hannah Fernando, and Dana Eardley. 2016. {``The Social
Lab Classroom: Wrestling with---and Learning from---Sustainability
Challenges.''} \emph{Sustainability: Science, Practice, and Policy} 12
(March): 76--87. \url{https://doi.org/10.1080/15487733.2016.11908155}.

\leavevmode\hypertarget{ref-Leonard1995}{}%
Leonard-Barton, Dorothy. 1995. \emph{Wellsprings of Knowledge: Building
and Sustaining the Sources of Innovation}. Harvard Business School.
\url{http://id.lib.harvard.edu/alma/990060514740203941/catalog}.

\leavevmode\hypertarget{ref-Lewis2020}{}%
Lewis, Jenny M. 2020. {``The Limits of Policy Labs: Characteristics,
Opportunities and Constraints.''} \emph{Policy Design and Practice},
December, 1--10. \url{https://doi.org/10.1080/25741292.2020.1859077}.

\leavevmode\hypertarget{ref-Lewis2005}{}%
Lewis, Michael, and James Moultrie. 2005. {``The Organizational
Innovation Laboratory.''} \emph{Creativity and Innovation Management} 14
(March): 73--83. \url{https://doi.org/10.1111/j.1467-8691.2005.00327.x}.

\leavevmode\hypertarget{ref-Magadley2009}{}%
Magadley, Wissam, and Kamal Birdi. 2009. {``Innovation Labs: An
Examination into the Use of Physical Spaces to Enhance Organizational
Creativity.''} \emph{Creativity and Innovation Management} 18
(December): 315--25.
\url{https://doi.org/10.1111/j.1467-8691.2009.00540.x}.

\leavevmode\hypertarget{ref-McGann2019}{}%
McGann, Michael, Tamas Wells, and Emma Blomkamp. 2019. {``Innovation
Labs and Co-Production in Public Problem Solving.''} \emph{Public
Management Review}, December, 1--20.
\url{https://doi.org/10.1080/14719037.2019.1699946}.

\leavevmode\hypertarget{ref-Memon2018}{}%
Memon, Atia Bano, Kyrill Meyer, Michael Thieme, and Lars-Peter Meyer.
2018. {``Inter-InnoLab Collaboration: An Investigation of the Diversity
and Interconnection Among Innovation Laboratories.''} \emph{Journal of
Engineering and Technology Management} 47 (January): 1--21.
\url{https://doi.org/10.1016/j.jengtecman.2017.11.003}.

\leavevmode\hypertarget{ref-Moulaert2014}{}%
Moulaert, Frank, Diana MacCallum, Abdelillah Hamdouch, and Abid Mehmood.
2014. \emph{The International Handbook on Social Innovation: Collective
Action, Social Learning and Transdisciplinary Research}. Illustrated,
reprint. Edward Elgar.
\url{https://books.google.fr/books?id=WWqdoAEACAAJ}.

\leavevmode\hypertarget{ref-Mulder2014}{}%
Mulder, Martin. 2014. {``Conceptions of Professional Competence.''} In
\emph{International Handbook of Research in Professional and
Practice-Based Learning}, edited by Stephen Billett, Christian Harteis,
and Hans Gruber, 107--37. Springer, Dordrecht.
\url{https://doi.org/10.1007/978-94-017-8902-8_5}.

\leavevmode\hypertarget{ref-Mulgan2006}{}%
Mulgan, Geoff. 2006. {``The Process of Social Innovation.''}
\emph{Innovations: Technology, Governance, Globalization} 1 (April):
145--62. \url{https://doi.org/10.1162/itgg.2006.1.2.145}.

\leavevmode\hypertarget{ref-Murray2010}{}%
Murray, Robin, Julie Caulier-Grice, and Geoff Mulgan. 2010. \emph{The
Open Book of Social Innovation}. National endowment for science,
technology; the art London.
\url{https://youngfoundation.org/publications/the-open-book-of-social-innovation/}.

\leavevmode\hypertarget{ref-Nilsson2015}{}%
Nilsson, Warren, Francois Bonnici, and Eliada W. Griffin EL. 2015.
{``The Social Innovation Lab: An Experiment in the Pedagogy of
Institutional Work.''} In \emph{The Business of Social and Environmental
Innovation: New Frontiers in Africa}, 201--12. Springer International
Publishing. \url{https://doi.org/10.1007/978-3-319-04051-6_11}.

\leavevmode\hypertarget{ref-Nystrom2014}{}%
Nyström, Anna-Greta, Seppo Leminen, Mika Westerlund, and Mika
Kortelainen. 2014. {``Actor Roles and Role Patterns Influencing
Innovation in Living Labs.''} \emph{Industrial Marketing Management} 43
(April): 483--95. \url{https://doi.org/10.1016/j.indmarman.2013.12.016}.

\leavevmode\hypertarget{ref-Osorio2019}{}%
Osorio, Ferney, Laurent Dupont, Mauricio Camargo, Pedro Palominos, José
Ismael Peña, and Miguel Alfaro. 2019. {``Design and Management of
Innovation Laboratories: Toward a Performance Assessment Tool.''}
\emph{Creativity and Innovation Management} 28 (January): 82--100.
\url{https://doi.org/10.1111/caim.12301}.

\leavevmode\hypertarget{ref-Osorio2019-A}{}%
Osorio, Ferney, Laurent Dupont, Mauricio Camargo, and Jose Ismael Pena.
2019. {``Constellation of Innovation Laboratories: A Scientific
Outlook.''} In \emph{2019 IEEE International Conference on Engineering,
Technology and Innovation (ICE/ITMC)}, 1--10. IEEE.
\url{https://doi.org/10.1109/ICE.2019.8792816}.

\leavevmode\hypertarget{ref-Osorio2020}{}%
Osorio, Ferney, Laurent Dupont, Mauricio Camargo, Carlos Sandoval, and
José Ismael Peña. 2020. {``Shaping a Public Innovation Laboratory in
Bogota: Learning Through Time, Space and Stakeholders.''} \emph{Journal
of Innovation Economics \& Management} 31: 69--100.
\url{https://doi.org/10.3917/jie.pr1.0066}.

\leavevmode\hypertarget{ref-Petersen2021}{}%
Petersen, Il haam, and Glenda Kruss. 2021. {``Universities as Change
Agents in Resource-Poor Local Settings: An Empirically Grounded Typology
of Engagement Models.''} \emph{Technological Forecasting and Social
Change} 167 (June): 120693.
\url{https://doi.org/10.1016/j.techfore.2021.120693}.

\leavevmode\hypertarget{ref-Podmetina2015}{}%
Podmetina, Daria, Joachim Hafkesbrink, Roman Teplov, Justyna Dabrowska,
and Monika Petraite. 2015. {``What Skills and Competences Are Required
to Implement Open Innovation?''} In \emph{ISPIM Conference Proceedings},
1--20. The International Society for Professional Innovation Management
(ISPIM).
\url{https://search.proquest.com/docview/1780140077/fulltext/11860BF316DB4B68PQ/1?accountid=14211}.

\leavevmode\hypertarget{ref-Podmetina2018}{}%
Podmetina, Daria, Klas Eric Soderquist, Monika Petraite, and Roman
Teplov. 2018. {``Developing a Competency Model for Open Innovation: From
the Individual to the Organisational Level.''} \emph{Management
Decision}, May. \url{https://doi.org/10.1108/MD-04-2017-0445}.

\leavevmode\hypertarget{ref-Puttick2014-Teams}{}%
Puttick, Ruth, Peter Baeck, and Philip Colligan. 2014. \emph{I-Teams:
The Teams and Funds Making Innovation Happen in Governments Around the
World}. Nesta.
\url{https://www.nesta.org.uk/report/i-teams-the-teams-and-funds-making-innovation-happen-in-governments-around-the-world/}.

\leavevmode\hypertarget{ref-Rayna2019}{}%
Rayna, Thierry, and Ludmila Striukova. 2019. {``Open Social Innovation
Dynamics and Impact: Exploratory Study of a Fab Lab Network.''} \emph{R
and D Management} 49 (June): 383--95.
\url{https://doi.org/10.1111/radm.12376}.

\leavevmode\hypertarget{ref-Rayna2021}{}%
---------. 2021. {``Fostering Skills for the 21st Century: The Role of
Fab Labs and Makerspaces.''} \emph{Technological Forecasting and Social
Change} 164 (March): 120391.
\url{https://doi.org/10.1016/j.techfore.2020.120391}.

\leavevmode\hypertarget{ref-Roberts1982}{}%
Roberts, Edward B., and Alan R. Fusfeld. 1982. {``Critical Functions:
Needed Roles in the Innovation Process.''} In \emph{Carrer Issues in
Human Resource Management}, edited by Ralph Kats, 182--207.
Prentice-Hall.
\url{https://books.google.fr/books/about/Career_issues_in_human_resource_manageme.html?id=4CYUAQAAMAAJ\&redir_esc=y}.

\leavevmode\hypertarget{ref-Sandberg2000}{}%
Sandberg, Jörgen. 2000. {``Understanding Human Competence at Work: An
Interpretative Approach.''} \emph{Academy of Management Journal} 43
(February): 9--25. \url{https://doi.org/10.5465/1556383}.

\leavevmode\hypertarget{ref-Strasser2019}{}%
Strasser, Tim, Joop de Kraker, and René Kemp. 2019. {``Developing the
Transformative Capacity of Social Innovation Through Learning: A
Conceptual Framework and Research Agenda for the Roles of Network
Leadership.''} \emph{Sustainability (Switzerland)} 11 (March).
\url{https://doi.org/10.3390/su11051304}.

\leavevmode\hypertarget{ref-Taber2018}{}%
Taber, Keith S. 2018. {``The Use of Cronbach's Alpha When Developing and
Reporting Research Instruments in Science Education.''} \emph{Research
in Science Education} 48 (December): 1273--96.
\url{https://doi.org/10.1007/s11165-016-9602-2}.

\leavevmode\hypertarget{ref-Timeus2018}{}%
Timeus, K, and M Gascó. 2018. {``Increasing Innovation Capacity in City
Governments: Do Innovation Labs Make a Difference?''} \emph{Journal of
Urban Affairs} 40: 992--1008.
\url{https://doi.org/10.1080/07352166.2018.1431049}.

\leavevmode\hypertarget{ref-Timmermans2020}{}%
Timmermans, Job, Vincent Blok, Robert Braun, Renate Wesselink, and
Rasmus Øjvind Nielsen. 2020. {``Social Labs as an Inclusive Methodology
to Implement and Study Social Change: The Case of Responsible Research
and Innovation.''} \emph{Journal of Responsible Innovation} 7
(September): 410--26.
\url{https://doi.org/10.1080/23299460.2020.1787751}.

\leavevmode\hypertarget{ref-RezaeeVessal2021}{}%
Vessal, Saeedeh Rezaee, Judith Partouche-Sebban, Veronica Scuotto, and
Adnane Maalaoui. 2021. {``Overcoming Stressful Life Events at
Do-It-Yourself (DIY) Laboratories. A New Trailblazing Career for
Disadvantaged Entrepreneurs.''} \emph{Technological Forecasting and
Social Change} 164 (March): 120506.
\url{https://doi.org/10.1016/j.techfore.2020.120506}.

\leavevmode\hypertarget{ref-Wascher2018}{}%
Wascher, Eva, Florian Hebel, Katharina Schrot, and Jürgen Schultze.
2018. {``Social Innovation Labs - a Starting Point for Social
Innovation.''} TU Dortmund University.
\url{https://kommunen-innovativ.de/sites/default/files/kosi-lab_report_social_innovation_labs_final.pdf}.

\leavevmode\hypertarget{ref-Wascher2019}{}%
Wascher, Eva, Christoph Kaletka, and Jürgen Schultze. 2019. {``Social
Innovation Labs - a Seedbed for Social Innovation.''} In \emph{Atlas of
Social Innovation: 2nd Volume - A World of New Practices}, 136--38.
\url{https://www.socialinnovationatlas.net/articles/}.

\leavevmode\hypertarget{ref-Westley2015}{}%
Westley, F., S. Laban, C. Rose, K. McGowan, K. Robinson, O. Tjornbo, and
M. Tovey. 2015. {``Social Innovation Lab Guide.''} Waterloo Institute
for Social Innovation; Resilience: Waterloo.
\url{https://uwaterloo.ca/waterloo-institute-for-social-innovation-and-resilience/projects/social-innovation-lab-guide}.

\leavevmode\hypertarget{ref-Zivkovic2018}{}%
Zivkovic, Sharon. 2018. {``Systemic Innovation Labs: A Lab for Wicked
Problems.''} \emph{Social Enterprise Journal} 14 (August): 348--66.
\url{https://doi.org/10.1108/SEJ-04-2018-0036}.

\end{CSLReferences}

\bibliography{CAIM2021.bib}


\end{document}